### 2019-03-13 paper/sex_selection_spacing-ver2.tex
Write up of high education results

Women with eight or more years of education are those expected to have the lowest 
fertility and the highest use of sex selection based on the prior literature.
Figure \ref{fig:bootstrap_high} shows that not only do this group, indeed, have lower 
fertility than the two prior groups, it is also falling rapidly over time.
Of women with one birth, the probability of having a second birth when the first-born
is a son is only around 70 percent for urban women and around 80 percent for rural women,
and for those with a first-born daughter it is around 80 percent for urban women and 90 
percent for rural women.
The falling fertility is even more clear for the third and fourth spells when the woman
has given birth to at least one son.
At the beginning of the sample, urban women with at least one boy had a close to 75
percent likelihood of having a birth, but that has fallen to only around 25 percent.
However, son preference still shows clearly in the parity progression probability for 
women without a son, with the decline for the third and fourth spell substantially
slower for both urban and rural women without a son than for those with at least one son.

Women with eight or more year of education mostly followed 
the traditional spacing pattern before sex selection became 
available as shown in Figure \ref{fig:bootstrap_high}.
Compared to the two other education groups there are,
however, more situations where the pattern is less clear.
Part of the problem may be that there are few women in
this group for the 1972--1984 period, and, hence, only 
the results for the first and second spells in both urban 
and rural areas are likely reliable.

Birth spacing increases substantially over time, 
especially for urban women. 
The median duration for urban women increases between a half
and a full year, whereas the increase for rural women 
is between a quarter and half a year.
The spacing between marriage and the first birth is still the
exception, and for this group of women, there is a decline of
about a month for both urban and rural areas.

In urban areas, the increase in spacing with only girls is so 
substantial that women with only girls have longer spacing to the 
next birth than any of the other sex compositions from 1985 onward.%
\footnote{
The exception is the second spell for the 1985--1994 period, where women
with one girl have 0.1 months shorter spacing the women with one boy.
}
This reversal in the standard spacing pattern is substantial, 
especially for the third and fourth spells.
For the latest period, the difference between only girls than the
other sex compositions is between 4.8 and 9.1 months for the third
and fourth spell.
Furthermore, the differences are statistically significant, except
for only boys for the fourth spell because few women have a fourth
birth if they have three boys already.

Rural women also show a reversal of the typical spacing pattern,
although the changes are less pronounced than for urban women.
The third spell shows that spacing for women with only girls become
increasing longer compared to the other two sex compositions.
Most of the differences are statistically significant, although they are 
smaller than for urban women.
For the second spell, there is little difference in spacing for
women with one girl and women with one boy.

The predicted sex ratios at the end of the spells show that
the reversal in spacing patterns is not the result of a declining son
preference but instead corresponds to a substantially more male-biased 
sex ratio.
For urban women, the predicted sex ratio with only girls 
is consistently above 60\% boys for the third and fourth spells.
For rural women, the predicted sex ratios are lower than for urban
women but still substantial and statistically significant at close 
to 60\%.
The high predicted sex ratios are presumably the result of increased 
use of sex selection.

>>>>

Women with eight or more years of education are those expected to have the lowest 
fertility and the highest use of sex selection based on the prior literature.
Figure \ref{fig:bootstrap_high} shows that not only do this group, indeed, have lower 
fertility than the two prior groups, it is also falling rapidly over time.
Of women with one birth, the probability of having a second birth when the first-born
is a son is only around 70 percent for urban women and around 80 percent for rural women,
and for those with a first-born daughter it is around 80 percent for urban women and 90 
percent for rural women.

The rapid decline in fertility for the most educated women is even more clear for the 
third and fourth spells when the woman has given birth to at least one son.
At the beginning of the sample, urban women with at least one boy had a close to 75
percent likelihood of having a birth, but that has fallen to only around 25 percent.
However, son preference still shows clearly in the parity progression probability for 
women without a son, with the decline for the third and fourth spell substantially
slower for both urban and rural women without a son than for those with at least one son.
The falling fertility means that relatively few women makes it to the fourth spell,
which, therefore, has more noise than the other spells.

Average birth spacing for the second spell is almost identically across sex compositions 
for both urban and rural women, and increased by around five months for rural women and
by almost ten months for urban women over the four decades.
These similarities hides, however, what would have been a standard son preference spacing 
pattern with significantly shorter spacing after a first-born girl than after a first-born 
boy if there had not been a substantial amount of sex selection.
For both of the last two periods the predicted sex ratio for women with a daughter as 
their first-born child is above 55 percent for both urban and rural areas.

The tremendous impact that sex selection can have on birth spacing is illustrated 
particularly vividly by the changes over time for the third spell.
In the 1972-1984 period the predicted average birth intervals for the three possible sex 
compositions were less than two months apart.
For the later periods, however, the highest predicted average spacing was for women
with only daughters with most of the differences to the other sex compositions 
statistically significant.
In urban areas the predicted average birth interval for a woman with only daughters 
increased by almost a full year from 26.8 to 38.3 months.
What is more, this increase is likely an underestimate.
The sex ratios for the 1972-84 period are substantially higher than the natural sex ratio 
despite sex selection not being widely available until around 1985.
If this early elevated sex ratio is because of recall error it means that the actual birth 
intervals would have been substantially shorter than the ones predicted here and the
increase over time larger.

The reversal pattern also shows for the fourth spell for urban women, although not for
rural women.
An important caveat is that because of the falling fertility relatively few women make
it to the fourth spell, and even if they do we do not observe a birth by the time of the
survey.
Of the 11,886 women across urban and rural areas observed for the fourth spell for the 
2005-2016 period more than 70 percent did not have a birth at the time of the survey.
Even though the censoring is less prevalent for the earlier periods this is countered by
those samples being substantially smaller.
Despite this, the pattern for the fourth spell is similar to that of the third spell.

The predicted sex ratios at the end of the spells show that this reversal in spacing 
patterns is not the result of a declining son preference but instead corresponds to a 
substantially more male-biased sex ratio as use of sex selection spread.
For urban women, the predicted sex ratio with only girls is consistently above 60\% boys 
for the third and fourth spells.
For rural women, the predicted sex ratios are lower than for urban women but still 
substantial and statistically significant at close to 60\%.
